---
title: "Final-Clean"
author: "Kelsey King"
date: "2024-02-17"
output: html_document
---
# Introduction
The following code pertains to the manuscript "Fecundity without nectar is insufficient for persistence of a blue butterfly" by Kelsey C. King and Cheryl B. Schultz.

Loading various libraries
```{r}
library(ggplot2)
library(ggpubr)
library(grid)
library(dplyr)
library(lme4)
library(MASS)
library(car)
library(emmeans)
library(forcats)
library(glm.predict)
#If you need to install bootpredictlme4 use code in hashtag below
#require("remotes")
#remotes::install_github("RemkoDuursma/bootpredictlme4")
library(bootpredictlme4)

#from Dealing with quasi- models in R by Ben Bolker (2023) https://cran.r-project.org/web/packages/bbmle/vignettes/quasi.pdf
#function for glms
dfun <- function(object) {
with(object,sum((weights * residuals^2)[weights > 0])/df.residual)
}
#from Ben Bolkder's Mixed Model Extensions 
#https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-for-overdispersioncomputing-overdispersion-factor
#function for testing for overdispersion in glmer models
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

# Sugar Experiment
Please reference the manuscript for details on the treatments.
```{r}
#summary data of adults
exp.adults=read.csv("./Adults2019_Puget.csv",stringsAsFactors = TRUE)

#daily egg laying of adults
laid.eggs=read.csv("./DailyFec2019_Puget.csv",stringsAsFactors = TRUE)

#removing individuals that do not meet criteria and loading other data
exp.adults=exp.adults[!(exp.adults$Indi=="A5"|exp.adults$Indi=="A10"),]
laid.eggs=laid.eggs[!(exp.adults$Indi=="A5"|exp.adults$Indi=="A10"),]

#dissection data with counts of unlaid eggs
dissect=read.csv("./Dissect2019_Puget.csv",stringsAsFactors = TRUE)
dissect=dissect[!(dissect$Indi=="A5"|dissect$Indi=="A10"|dissect$Indi=="A9"),]

#weight measurements of adults throughout experiment
weight=read.csv("./Weights2019_Puget.csv",stringsAsFactors = TRUE)
weight=weight[!(weight$Indi=="A5"|weight$Indi=="A10"),]

#details needed throughout this experiment
#Experimental title, colors, treatment labels etc.
#days of interest in daily fecundity and weight data
exp.title=c("Sugar")
exp.days1=c(1,3,5,10,18)
exp.days2=c(1,3,5,11,17)
t.names=c("A", "R", "W")
t.colors=c("#27eb5a","#eb5a27","#5a27eb")
t.labs=c('Ad Libitum', 'Restricted', 'Water Only')
```

## Total Fecundity
We used the same process to evaluate the Sugar and Amino Acids experimental models as used below, and will not repeat that code below, for reference review each model here.
```{r}
#total fecundity is the sum of all eggs laid by each female
tot.fec=glm(Total.Fecundity ~ Treat, data=exp.adults, family="poisson")
tot.fec.q=glm(Total.Fecundity ~ Treat, data=exp.adults, family="quasipoisson")
summary(tot.fec.q)
#overdispersed

#other model fits, testing AIC
tot.fec.nb=glm.nb(Total.Fecundity ~ Treat, data=exp.adults)
bbmle::qAIC(tot.fec.nb, dispersion= dfun(tot.fec.nb), nobs=nrow(exp.adults))
#qAIC here uses the dispersion calculated to adjust the Poisson model, e.g. what is the AIC if we adjust the Poisson model to be a quasi-Poisson
bbmle::qAIC(tot.fec,dispersion=dfun(tot.fec),nobs=nrow(exp.adults))
#quasipoisson performing better than glm.nb

tot.fec=tot.fec.q
plot(tot.fec)

#is treatment significant
Anova(tot.fec)

#which treatments differ from each other
marginal = emmeans(tot.fec,~ Treat)
pairs(marginal, adjust="tukey")

#prediction interval for summary in table
tot.pred=glm.predict::predicts(tot.fec, "F")
```

## Longevity
```{r}
#longevity is the total number of days an adult lived from capture to death
longevity=glm(Days.in.Trial ~ Treat, data=exp.adults, family='poisson')
long.q=glm(Days.in.Trial ~ Treat, data=exp.adults, family='quasipoisson')
summary(long.q)
#overdispersed

long.nb=glm.nb(Days.in.Trial ~ Treat, data=exp.adults)
qAIC(long.nb, dispersion=dfun(long.nb), nobs=nrow(exp.adults))
qAICc(longevity,dispersion=dfun(longevity),nobs=nrow(exp.adults))
#quasipoisson performing better than glm.nb

longevity=long.q
plot(longevity)
#treatment significant
Anova(longevity)

#which treatments differ
marginal = emmeans(longevity,~ Treat)
pairs(marginal, adjust="tukey")

#prediction interval for summary in table, means are used in population model
lfspn.pred=glm.predict::predicts(longevity, "F",type="bootstrap",conf.int = 0.95,doPar=TRUE,sim.count=1000)
```

## Daily Fecundity
```{r}
#Poisson
daily.fec.init=glmer(Total.eggs ~ Treat*Day + (1 | Indi), data=laid.eggs, family='poisson')
#test for overdispersion (see intro.)
overdisp_fun(daily.fec.init)

#logistic Poisson
laid.eggs$ID.egg=rownames(laid.eggs)
daily.fec.logPoi=glmer(Total.eggs ~ Treat*Day + (1 | Indi)+ (1|ID.egg), data=laid.eggs, family='poisson', control = glmerControl(optimizer = "bobyqa"))
overdisp_fun(daily.fec)

#Negative Binomial
daily.fec.nb=glmer.nb(Total.eggs ~ Treat*Day + (1 | Indi), data=laid.eggs)
overdisp_fun(daily.fec.nb)

#model comparison
AIC(daily.fec.logPoi)
AIC(daily.fec.nb)

#selected negative binomial
daily.fec=daily.fec.nb

#evaluate fit
plot(daily.fec,residuals(.) ~log(fitted(.)))
plot(daily.fec,Indi~resid(.,type="pearson"))

#significant model coefficients
Anova(daily.fec)

#significant treatment differences
#must be by day based on model structure
em.day.fec=emmeans(daily.fec,~Treat|Day, at=list(Day=exp.days1),type="response")
pwpm(em.day.fec)
plot(em.day.fec,comparisons = TRUE)
```

This code will take a while as it is creating the prediction interval for the data visualization of daily fecundity. The prediction interval mean is also used in the population model.

```{r}
treats=unique(laid.eggs$Treat)
Treat=rep(treats,max(exp.adults$Days.in.Trial))
Day=sort(rep(c(1:max(exp.adults$Days.in.Trial)),length(treats)))
day.fec.pred=data.frame(Treat,Day)

#get the predicted mean daily egg lay across all individuals by
#treatment using bootpredictlme4
#his will be time consuming

PR=predict(daily.fec,newdata=day.fec.pred,re.form=NA,se.fit=TRUE,nsim=1000,type="response")

#putting results into data frame
day.fec.pred$DailyEgg=PR$fit
ci.pred=t(PR$ci.fit)
day.fec.pred$lower=ci.pred[,1]
day.fec.pred$upper=ci.pred[,2]
day.fec.pred$se.boot=PR$se.boot
day.fec.pred$Treat=as.character(day.fec.pred$Treat)

#Removing the predictions for the days after which 97.5% of butterflies would have died as estimated in the longevity model prediction
day.results=data.frame()
treats=as.vector(unique(day.fec.pred$Treat))

for(ea.treat in 1:length(treats)){
  cur.treat=treats[ea.treat]
  cur.dat=day.fec.pred[day.fec.pred$Treat==cur.treat,]
  cur.lfspn=lfspn.pred[lfspn.pred$Treat==cur.treat,]
  comp.res= cur.dat[!(cur.dat$Day > ceiling(cur.lfspn$upper)),]
  day.results=rbind(day.results,comp.res)
}
```

Plotting the prediction interval for this experiment, will be placed alongside the other treatment later.

```{r}
#making the results into a figure
fig.df=ggplot(aes(Day,DailyEgg,fill=Treat),data=day.results)+geom_errorbar(aes(ymin=lower,ymax=upper), linewidth=0.5,width=0,position = position_dodge(width=1), color="gray40")+
  geom_point(size=3,shape=21,position=position_dodge(width=1))+
  scale_fill_manual(name="", breaks=t.names, values=t.colors, labels=function(x)stringr::str_wrap(t.labs,8))+scale_color_manual(name="", breaks=t.names, values=t.colors, guide="none")+
  #theme components
  xlab("")+ylab("")+theme_pubr(base_size = 14)+ylim(0,60)+xlim(0,26)+ theme(legend.position = c(0.75, .72), legend.background = element_blank(),legend.text=element_text(size=11),axis.line.x.bottom=element_line(linewidth =1),axis.line.y.left=element_line(linewidth =1),axis.ticks=element_line(linewidth=1), plot.margin = margin(t = 0,r = 0,b = 0,  l = 2)) +labs(x = NULL, y=NULL)+ annotate("text", x = 26, y = 60, label = "a", size=6)
#check figure
fig.df

```

## Unlaid Eggs
```{r}
dissect.mod=glm(Lost.Fecundity~Treat, data=dissect, family="poisson")

dfun(dissect.mod)
#overdispersed

dissect.nb=glm.nb(Lost.Fecundity~Treat, data=dissect)

#qAIC from bbmle
bbmle::qAIC(dissect.nb, dispersion= dfun(dissect.nb), nobs=nrow(dissect))
#this calculates the AIC of the quasipoisson, given the Poisson model and its overdispersion
bbmle::qAIC(dissect.mod,dispersion=dfun(dissect.mod),nobs=nrow(dissect))

#quasipoisson perform better than glm.nb
dissect.modq=glm(Lost.Fecundity~Treat, data=dissect, family="quasipoisson")
plot(dissect.mod)

#treatment significant
Anova(dissect.mod)

#which treatments differ
em.dissect=emmeans(dissect.mod,~Treat,type="response")
pairs(em.dissect)
plot(em.dissect,comparisons = TRUE)

#prediction interval for tabel
lost.pred=glm.predict::predicts(dissect.mod, "F")

```

## Weight

```{r}
hist(weight$Wt)
#normal distribution
weight.mod=lmer(Wt~Day*Treat +(1 | Indi), data=weight)
summary(weight.mod)

plot(weight.mod,residuals(.) ~log(fitted(.)))

#coefficient significance
Anova(weight.mod)

#which treatments vary and when
#day is required because of the structure of the model
em.weight=emmeans(weight.mod,~Treat|Day, at=list(Day=exp.days2),type="response")
pwpm(em.weight)
plot(em.weight,comparisons = TRUE)

```

## Table Part 1
Creating a table for prediction intervals.
```{r}
longevity=vector()
for(each in 1:length(t.labs)){
  cur.dat=lfspn.pred[each,]
  pasted=paste0(round(cur.dat$mean,0)," (", round(cur.dat$lower,0),", ",round(cur.dat$upper,0),")")
  longevity[each]=pasted
}

total.fec=vector()
for(each in 1:length(t.labs)){
  cur.dat=tot.pred[each,]
  pasted=paste0(round(cur.dat$mean,0)," (", round(cur.dat$lower,0),", ",round(cur.dat$upper,0),")")
  total.fec[each]=pasted
}

lost.vals=vector()
for(each in 1:length(t.labs)){
  cur.dat=lost.pred[each,]
  pasted=paste0(round(cur.dat$mean,0)," (", round(cur.dat$lower,0),", ",round(cur.dat$upper,0),")")
  lost.vals[each]=pasted}

nect.nut=matrix(c(total.fec,longevity,lost.vals), ncol = 3,byrow=F)


rownames(nect.nut)=t.labs
colnames(nect.nut)=c("Fecundity (95% CI)", "Longevity (95% CI)", paste("Unlaid Eggs","(95% CI)"))

knitr::kable(nect.nut)
```


# Sugar and Amino Acid Experiment
Model selection underwent the same process, but for the purposes of not repeating code is not done here.
```{r}
#loading data and removing individuals that did not meet criteria
laid.eggs.sAA=read.csv("./DailyFec2020.csv",stringsAsFactors = TRUE)
exp.adults.sAA=read.csv("./Adults2020.csv",stringsAsFactors = TRUE)
weight.sAA=read.csv("./Weight2020.csv",stringsAsFactors = TRUE)
dissect.sAA=read.csv("./Dissect2020.csv",stringsAsFactors = TRUE)

#removing individuals see methods for details
exp.adults.sAA=exp.adults.sAA[!(exp.adults.sAA$Indi=="L02"|exp.adults.sAA$Indi=="L08"),]
laid.eggs.sAA=laid.eggs.sAA[!(laid.eggs.sAA$Indi=="L02"|laid.eggs.sAA$Indi=="L08"),]
weight.sAA=weight.sAA[!(weight.sAA$Indi=="L02"|weight.sAA$Indi=="L08"),]
dissect.sAA=dissect.sAA[!(dissect.sAA$Indi=="L02"|dissect.sAA$Indi=="L08"),]

#title for saving figures and days of interest in weight data
exp.title.sAA=c("SugarAminoAcids")
exp.days1.sAA=c(1,10,18)
exp.days2.sAA=c(1,14,15,28)
t.names.sAA=c("B","F", "L", "W")
t.colors.sAA=c("#27eb5a","#ebbc27","#eb27b8","#2756eb")
t.labs.sAA= c("Flower & Lupine", 'Flower',"Lupine", 'Water Only')
```

## Total Fecundity
```{r}
#model
tot.fec.sAA=glm(Total.Fecundity ~ Treat, data=exp.adults.sAA, family="quasipoisson")
summary(tot.fec.sAA)
plot(tot.fec.sAA)

#treatment effects
Anova(tot.fec.sAA)
marginal = emmeans(tot.fec.sAA,~ Treat)
pairs(marginal, adjust="tukey")

#prediction interval
tot.pred.sAA=glm.predict::predicts(tot.fec.sAA, "F")

```
## Longevity
```{r}
#model
longevity.sAA=glm(Days.in.Trial ~ Treat, data=exp.adults.sAA, family='quasipoisson')
summary(longevity.sAA)
plot(longevity.sAA)

#treatment effects
Anova(longevity.sAA)
marginal = emmeans(longevity.sAA,~ Treat)
pairs(marginal, adjust="tukey")

#prediction interval
lfspn.pred.sAA=glm.predict::predicts(longevity.sAA, "F",type="bootstrap",conf.int = 0.95,doPar=TRUE,sim.count=1000)
```

## Daily Fecundity
```{r}
#model
daily.fec.nb=glmer.nb(Total.eggs ~ Treat*Day + (1 | Indi), data=laid.eggs.sAA)
overdisp_fun(daily.fec.nb)

daily.fec.sAA=daily.fec.nb
plot(daily.fec.sAA,residuals(.) ~log(fitted(.)))
plot(daily.fec.sAA,Indi~resid(.,type="pearson"))

#treatment effects
Anova(daily.fec.sAA)
em.day.fec=emmeans(daily.fec.sAA,~Treat|Day, at=list(Day=exp.days1.sAA),type="response")
pwpm(em.day.fec)
plot(em.day.fec,comparisons = TRUE)
```

This code will take a while as it is creating the prediction interval for the data visualization of daily fecundity. The prediction interval mean is also used in the population model.
```{r}
treats.sAA=unique(laid.eggs.sAA$Treat)
Treat=rep(treats.sAA,max(exp.adults.sAA$Days.in.Trial))
Day=sort(rep(c(1:max(exp.adults.sAA$Days.in.Trial)),length(treats.sAA)))
day.fec.pred.sAA=data.frame(Treat,Day)

#get the predicted mean daily egg lay across all individuals by
#treatment using bootpredictlme4
PR.sAA=predict(daily.fec.sAA,newdata=day.fec.pred.sAA,re.form=NA,se.fit=TRUE,nsim=1000,type="response")

#putting results into data frame
day.fec.pred.sAA$DailyEgg=PR.sAA$fit
ci.pred=t(PR.sAA$ci.fit)
day.fec.pred.sAA$lower=ci.pred[,1]
day.fec.pred.sAA$upper=ci.pred[,2]
day.fec.pred.sAA$se.boot=PR.sAA$se.boot
day.fec.pred.sAA$Treat=as.character(day.fec.pred.sAA$Treat)

#Removing the predictions for the days after which 97.5% of butterflies would have died as estimated in the longevity model prediction
day.results.sAA=data.frame()
for(ea.treat in 1:length(treats.sAA)){
  cur.treat=as.character(treats.sAA)[ea.treat]
  cur.dat=day.fec.pred.sAA[day.fec.pred.sAA$Treat==cur.treat,]
  cur.lfspn=lfspn.pred.sAA[lfspn.pred.sAA$Treat==cur.treat,]
  comp.res= cur.dat[!(cur.dat$Day > ceiling(cur.lfspn$upper)),]
  day.results.sAA= rbind(day.results.sAA,comp.res)}
```


```{r}
#making the results into a figure and saving it.
fig.df.sAA=ggplot(aes(Day,DailyEgg,fill=Treat),data=day.results.sAA)+geom_errorbar(aes(ymin=lower,ymax=upper), linewidth=0.5,width=0,position = position_dodge(width=1), color="gray40")+
  geom_point(size=3,shape=21,position=position_dodge(width=1))+ scale_fill_manual(name="", breaks=t.names.sAA, values=t.colors.sAA, labels=function(x)stringr::str_wrap(t.labs.sAA,8))+scale_color_manual(name="", breaks=t.names.sAA, values= t.colors.sAA, guide= "none")+theme_pubr(base_size = 14)+ylim(0,60)+xlim(0,26)+ theme(axis.line.x.bottom= element_line(linewidth =1), axis.line.y.left= element_line(linewidth =1), legend.position = c(0.75, .7), legend.background= element_blank(),legend.text= element_text(size = 11),axis.ticks= element_line(linewidth = 1), plot.margin = margin(t = 0,r = 0,b = 0,  l = 2)) +labs(x = NULL, y=NULL)+ annotate("text", x = 26, y = 60, label = "b", size=6)

fig.df.sAA

fig=ggarrange(fig.df, fig.df.sAA + rremove("y.text"), ncol = 2, nrow = 1, widths =3.75, heights=3.4)
fig=annotate_figure(fig, left = textGrob("Eggs Laid", rot = 90, gp = gpar(cex = 1.3)),bottom = textGrob("Day", gp = gpar(cex = 1.3)))

fig
ggsave("Fig1.eps", dpi=600, height=3.4, width=7.5)
ggsave("Fig1.jpeg", dpi=600, height=3.4, width=7.5)
```

## Unlaid Eggs
```{r}
#model
dissect.sAA.mod=glm(Lost.Fecundity~Treat, data=dissect.sAA, family="quasipoisson")
summary(dissect.sAA.mod)
plot(dissect.sAA.mod)

#treatment effects
Anova(dissect.sAA.mod)
em.dissect.sAA=emmeans(dissect.sAA.mod,~Treat,type="response")
pairs(em.dissect.sAA)
plot(em.dissect.sAA,comparisons = TRUE)

#prediction interval
lost.pred.sAA=glm.predict::predicts(dissect.sAA.mod, "F")
```



## Weight
```{r}
#model
weight.sAA.mod=lmer(Wt~Day*Treat +(1 | Indi), data=weight.sAA)
summary(weight.sAA.mod)
plot(weight.sAA.mod,residuals(.) ~log(fitted(.)))

#treatment effects
Anova(weight.sAA.mod)
em.weight.sAA=emmeans(weight.sAA.mod,~Treat|Day, at=list(Day=exp.days2.sAA),type="response")
pwpm(em.weight.sAA)
plot(em.weight.sAA,comparisons = TRUE)
```

## Table Part 2
Pulling all the prediction intervals into a table for data visualization.
```{r}
longevity=vector()
for(each in 1:length(t.labs.sAA)){
  cur.dat=lfspn.pred.sAA[each,]
  pasted=paste0(round(cur.dat$mean,0)," (", round(cur.dat$lower,0),", ",round(cur.dat$upper,0),")")
  longevity[each]=pasted
}

total.fec=vector()
for(each in 1:length(t.labs.sAA)){
  cur.dat=tot.pred.sAA[each,]
  pasted=paste0(round(cur.dat$mean,0)," (", round(cur.dat$lower,0),", ",round(cur.dat$upper,0),")")
  total.fec[each]=pasted
}

lost.vals=vector()
for(each in 1:length(t.labs.sAA)){
  cur.dat=lost.pred.sAA[each,]
  pasted=paste0(round(cur.dat$mean,0)," (", round(cur.dat$lower,0),", ",round(cur.dat$upper,0),")")
  lost.vals[each]=pasted}
  
nect.nut2=matrix(c(total.fec,longevity,lost.vals), ncol = 3,byrow=F)

rownames(nect.nut2)=t.labs.sAA
colnames(nect.nut2)=c("Fecundity (95% CI)", "Longevity (95% CI)", paste("Unlaid Eggs","(95% CI)"))

nect.nut.comp=rbind(nect.nut,nect.nut2)
knitr::kable(nect.nut.comp)
```

# Population Model

## Parameters
```{r}
sims=10000 #number of simulations

surv.published=c(
  #Estimating overwintering survival constant (egg to post-diapause larvae)
#all estimates are from baseline management (annual mowing)

#Schultz and Crone 1998 one site one year
0.094, 

#Warchola et al. 2017 
0.013,0.021,0.021, 
#three years one site

# Schultz and Ferguson 2020
#egg to post-diapause for plots with >=25 eggs monitored (removed Hayfield, Main, North 2018)
#average for 9 sites across 2014-2018 n=42
0.029471822, 0.016504997,
0.015288148,0.012745439,
0.062705746,0.036459305,
0.058690404,0.01164928,
0.050309215,0.081929126,
0.020733541,0.019521844,
0.004065041,0.011091821,
0.021978022,0.012806637,
0.058170996,0.024644946,
0.017424242,0.166666667,
0.02605042,0.005952381,
0.018650794,0.036910458,
0.01010101,0.01,0.007142857,
0.005952381,0,0.020408163,
0.09,0.010416667,0.014285714,
0,0.063737374,0.008333333,0.005952381,
0,0,0.0259914,0,0.006944444) 

surv.est=mean(surv.published)
#estimate is constant


```

## Simulation
> In this simulation of population growth, we work at the population
> level. We assume that our estimate of longevity and daily fecundity
> represent the average of the population. We estimate longevity and
> daily fecundity annually to get a stochastic fecundity for that year.
> Additionally, we include annual environmental stochasticity to get a
> more realistic range of growth rates.

```{r}
####Simulation####

#automatic adjusting of the rnbinom function for a quasi poisson
rqpois <- function(n, mu, theta) {
  rnbinom(n = n, mu = mu, size = mu/(theta-1))
}

#where the final results will be put
pop.bind1=data.frame()
pop.bind2=data.frame() #where the results will be put

#Experiment loop
for(ea.dat in 1:2){
#which experiment
if(ea.dat==1){
#sugar
  treats=c("W", "A", "R")
  theta.day=lme4:::getNBdisp(daily.fec)
  theta=dfun(longevity)
  lfspn.par=lfspn.pred[,c(1,4)]
  parm.data=merge(day.results[,c(1:3,6)],lfspn.par, by.x="Treat",by.y = "Treat")
  colnames(parm.data)=c("Treat","Day","dailyEgg","egg.se","daysAlive")
} else{
#Sugar and amino acids
  treats=as.character(treats.sAA)
  theta.day=lme4:::getNBdisp(daily.fec.sAA)
  theta=dfun(longevity)
  lfspn.par=lfspn.pred.sAA[,c(1,4)]
  parm.data=merge(day.results.sAA[,c(1:3,6)],lfspn.par, by.x="Treat",by.y = "Treat")
  colnames(parm.data)=c("Treat","Day","dailyEgg","egg.se","daysAlive")}
  
pop.bind=data.frame()
#treatment loop
for(ea.treat in 1:length(treats)){
  cur.treat=treats[ea.treat]
  cur.parm=parm.data[parm.data$Treat==cur.treat,]
  lam=rep("NA",sims)
#Simulation loop, repeating calculation across variation
for (ea.sim in 1:sims){
  #randomly estimate the average annual adult longevity parameter from the longevity model
  cur.lfspn=rqpois(1,cur.parm$daysAlive[1], theta)
  #adjust if lifespan is less than 1.
  if(cur.lfspn<1){days=1}else{days=round(cur.lfspn,0)}
  if(days>max(cur.parm$Day)){days=max(cur.parm$Day)}
  #fecundity matrix for each day alive 
  cur.fec=matrix(nrow=days,ncol=2)
  
#randomly estimate the average annual daily fecundity by the age (in days) of butterflies from the daily fecundity model
  #daily loop, length dependent on lifespan
for(ea.day in 1:days){
  number=vector()
  cur.day.parm=cur.parm[cur.parm$Day==paste(ea.day),]
  fec.day=cur.day.parm[,3]
  cur.fec[ea.day,1]=ea.day
  number=rnbinom(n=1, mu=fec.day, size=fec.day/(theta.day-1))
  #adjust if we get a bad random number
  if(is.na(number)|number<1){number=fec.day}else {number}
  cur.fec[ea.day,2]=number
}
  
#sum daily fecundity to get average population annual fecundity
fecundity=sum(cur.fec[,2])
  
#####Population simulation math#####
  lam[ea.sim]=fecundity*(1/2)*surv.est
#Population size is the result of the fecundity times the population size in the previous year, minus those individuals that died before reaching adulthood given the environmental variation.
  
  }#close simulations and move to next treatment
  #results dataset
  sim=c(1:sims)
  Treat=c(rep(cur.treat,sims))
  pop.matrix=data.frame(cbind(lam,sim,Treat))
  pop.bind=rbind(pop.bind,pop.matrix)

  #put the results into the dataset
  #start the next simulation
  }#close treatment and loop
if(ea.dat==1){
  #place data in each experiment dataset
  pop.bind1=pop.bind}else{
    pop.bind2=pop.bind}
}
#formatting the results
pop.bind1$lam=as.numeric(pop.bind1$lam)
pop.bind1$Treat=as.factor(pop.bind1$Treat)
pop.bind1$sim=as.factor(pop.bind1$sim)
pop.bind1$lam[is.nan(pop.bind1$lam)]=NA
pop.bind1=pop.bind1[!is.na(pop.bind1$lam),]

pop.bind2$lam=as.numeric(pop.bind2$lam)
pop.bind2$Treat=as.factor(pop.bind2$Treat)
pop.bind2$sim=as.factor(pop.bind2$sim)
pop.bind2$lam[is.nan(pop.bind2$lam)]=NA
pop.bind2=pop.bind2[!is.na(pop.bind2$lam),]

```
## Results
```{r}
#calculate the confidence interval for the lambdas
for(ea.dat in 1:2){
  #each experiment
  if(ea.dat==1){treats=c("W", "A", "R")
  pop.dat=pop.bind1}else{
    treats=treats.sAA
    pop.dat=pop.bind2}
  lam.res=data.frame()
for(ea.treat in 1:length(treats)){
  #each treatment in that experiment
  cur.treat=treats[ea.treat]
  cur.dat=pop.dat[pop.dat$Treat==cur.treat,]
  lower=mean(cur.dat$lam)-1.96*sd(cur.dat$lam)
  upper=mean(cur.dat$lam)+1.96*sd(cur.dat$lam)
   if(lower<0){lower=0}
  Treat=paste(cur.treat)
  est=mean(cur.dat$lam)
  comp.res=data.frame(cbind(Treat,est,lower,upper))
  lam.res=rbind(lam.res,comp.res)
}
if(ea.dat==1){
  #place in the dataset's correct df
  lam.results=lam.res
}else{
  lam2.results=lam.res}
}

#format each df
colnames(lam.results)=c("Treatment","mean","lower", "upper")
lam.results$mean=as.numeric(lam.results$mean)  
lam.results$upper=as.numeric(lam.results$upper)
lam.results$lower=as.numeric(lam.results$lower)
lam.results$Treat=as.factor(lam.results$Treat)

colnames(lam2.results)=c("Treatment","mean","lower", "upper")
lam2.results$mean=as.numeric(lam2.results$mean)
lam2.results$upper= as.numeric(lam2.results$upper)
lam2.results$lower= as.numeric(lam2.results$lower)
lam2.results$Treat= as.factor(lam2.results$Treat)
```

```{r}
#data visualization
fig=ggplot(aes(Treat,mean),data=lam.results)+ geom_errorbar(aes(ymin=lower,ymax=upper),data=lam.results, width=0,linewidth=0.5) +  geom_point(size=3.5,shape=21,aes(fill=Treat),data=lam.results)+geom_hline(yintercept =1,linewidth=0.7)+ scale_fill_manual(name="Treatment", values=t.colors,breaks=t.names,labels = t.labs)+ geom_hline(yintercept=1.55,color="gray2",linetype=2,linewidth=0.7)+ylab(expression(lambda))+xlab("Treatment") + ylim(0,5)+scale_x_discrete(labels=function(x)stringr::str_wrap(t.labs,8))+ theme_pubr(base_size=14) +theme(legend.position = "none")+ annotate("text", x = 3.4, y = 5, label = "a", size=6)+
  theme(axis.line.x.bottom= element_line(linewidth =1), axis.line.y.left= element_line(linewidth =1), legend.background= element_blank(),axis.text.x=element_text(size = 11),axis.ticks= element_line(linewidth = 1), plot.margin = margin(t = 0,r = 0,b = 0,  l = 2)) +labs(x = NULL, y=NULL)

fig2=ggplot(aes(Treat,mean),data=lam2.results)+ 
  geom_errorbar(aes(ymin=lower,ymax=upper),data=lam2.results, width=0,linewidth=0.5) +  
  geom_point(size=3.5,shape=21,aes(fill=Treat),data=lam2.results)+
  geom_hline(yintercept =1,linewidth=0.7)+ 
  scale_fill_manual(name="Treatment", values=t.colors.sAA,breaks=t.names.sAA,labels = t.labs.sAA)+ 
  geom_hline(yintercept=1.55,color="gray2",linetype=2,linewidth=0.7)+ ylim(0,5)+
  scale_x_discrete(labels=function(x)stringr::str_wrap(t.labs.sAA,8))+ 
  theme_pubr(base_size=14) +
  theme(legend.position = "none")+
  annotate("text", x = 4.2, y = 5, label = "b", size=6)+
  theme(axis.line.x.bottom= element_line(linewidth =1), axis.line.y.left= element_line(linewidth =1), legend.background= element_blank(),axis.text.x = element_text(size = 11),axis.ticks= element_line(linewidth = 1), plot.margin = margin(t = 0,r = 0,b = 0,  l = 2)) +labs(x = NULL, y=NULL)
#put the two figures together
figure=ggarrange(fig, fig2 + rremove("y.text"), ncol = 2, nrow = 1, widths = c(2.25, 2.75), heights=3)
figure=annotate_figure(figure, left = textGrob(expression(lambda), rot = 0, gp = gpar(cex = 1.75), vjust=-1),bottom = textGrob("Treatment", gp = gpar(cex = 1.3)))
#check figure
figure
#save figure
ggsave("Fig2.eps", figure, dpi=600, height=3, width=5)
ggsave("Fig2.jpeg", figure, dpi=600, height=3, width=5)
```
## Model Sensitivity
```{r}
#to get lambda 1.55 fecundity must be greater than
(1.55)/(surv.est*1/2)
#to get lambda 1 fecundity must be greater than
(1)/(surv.est*1/2)

#average water fec
w.sAA=34
w.sugar=68

#to get lambda 1.55 at avg water fecundity
(1.55*adults)/(w.sugar*(adults/2))
(1.55*adults)/(w.sAA*(adults/2))

#to get lambda 1 at avg water fecundity
(1*adults)/(w.sugar*(adults/2))
(1*adults)/(w.sAA*(adults/2))

#proportion of years where water fecundity could reach 1.55
length(surv.published[surv.published>(1.55*adults)/(w.sugar*(adults/2))])/ length(surv.published)

#proportion of years where water fecundity could reach 1
length(surv.published[surv.published>(1*adults)/(w.sugar*(adults/2))])/ length(surv.published)

```

Evaluating the sensitivity of the model to fecundity and larval survival. Given a certain fecundity, what is the larval survival required to be to get to the target population growth rates?

```{r}
#range of immature survivals
surv1.loop=seq(0.01,.1,0.005)
fec.155=vector()
#loop to calculate
for(each in 1:length(surv1.loop)){
  cur.s1=surv1.loop[each]
  fec=1.55/(1/2*cur.s1)
  fec.155[each]=fec
}
#visualize
fig=ggplot()+geom_line(aes(x=surv1.loop*100, y=fec.155))+xlab("Immature Survival (%)")+ ylab("Total Fecundity")+ geom_vline(xintercept=surv.est*100, color='grey40', lty=2, linewidth=1) +theme_pubr(base_size = 14)+ theme(axis.line.x.bottom= element_line(linewidth =1), axis.line.y.left= element_line(linewidth =1), axis.ticks= element_line(linewidth = 1), plot.margin = margin(t = 0,r = 5,b = 0,  l = 2))
#check figure
fig
#save figure
ggsave("Fig3.jpeg", fig, dpi = 600, height=3.4, width=3.4)
ggsave("Fig3.eps", fig, dpi = 600, height=3.4, width=3.4)
```
# References

Schultz CB, Crone EE (1998) Burning prairie to restore butterfly habitat: a modeling approach to management tradeoffs for the Fender’s blue. Restoration Ecology 6:244–252

Schultz CB, Ferguson JA (2020) Demographic costs and benefits of herbicide‐based restoration to enhance habitat for an endangered butterfly and a threatened plant. Restor Ecol 28:564--572. https://doi.org/10.1111/rec.13102

Warchola N, Crone EE, Schultz CB (2017) Balancing ecological costs and benefits of fire for population viability of disturbance-dependent butterflies. J Appl Ecol 55:800--809. https://doi.org/10.1111/1365-2664.12983